---
title: "Homework1"
output: pdf_document
date: "2022-09-04"
author: "Aidan Lytle"
email: "aflytle@uncg.edu"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1

```{r}
# make the variables for the problem

m = 10
n = 1000
k = 100


if (k >= n)
{
  print('error: k must be less than n.')
}

M = matrix(runif(n*m), nrow = n)
newM = matrix(0*1:m, nrow = 1)


for (i in 1:m) # iterate over rows
{

  MCheck = M[,i] # vector we're checking over: ith col of M
  
  for (j in 1:ceiling(k/(2*m)))
  {
    maxRowInd = which.max(MCheck)[1] #find max row
    newMRow = M[maxRowInd,] #assign row
    newM = rbind(newM, newMRow) #bind row to matrix
    M = M[-c(maxRowInd),] #delete row from data
    MCheck = MCheck[-maxRowInd] #delete index from search vector
    
  }
  
  MCheck = M[,i] #have to reassign outside loop in case anything broke
  
  for (j in 1:ceiling(k/(2*m)))
  {
    minRowInd = which.min(MCheck)[1]
    newMRow = M[minRowInd,]
    newM = rbind(newM, newMRow)
    M = M[-c(minRowInd),]
    MCheck = MCheck[-minRowInd]
  }
  
}
newM = newM[-1,] #delete initial row of zeroes




for (l in 2:m) #plots for comparison
{
  plot(M[,1],M[,l] )
  points(newM[,1],newM[,l],col = 'red')
    
}

```
It should be clear from these plots, and from the structure of the problem, that this algorithm is a way to construct a representative sample from a larger dataset. This is because the largest differences along any axis are what is represented, which (when sampled over all) produces a set which maintains the variation in the initial set, while lowering the dimensionality.

***

# Problem 3

```{r}
#library(ggplot2)

n = 100


M1 = matrix(runif(n*n), ncol = n) #n^2 random entries 0,1
M2 = matrix(runif(n*n, 0, 100), ncol = n) #range from 0,100
M3 = matrix(runif(n*n, 10, 2000), ncol = n) #10,2000
M4 = matrix(runif(n*n, -500, 0), ncol = n) #-500,0
MN1 = matrix(runif(n*n*n*n), ncol = n) #n^4


cltTest <- function(m)
{
  newm = 0*(1:ncol(m))
  for (i in 1:ncol(m))
  {
    newm[i] = mean(m[,i])
  }
  hist(newm, breaks = 25)
  #return(newm)
}

# calling function to produce the results of the CLT test
cltTest(M1)
cltTest(M2)
cltTest(M3)
cltTest(M4)
cltTest(MN1)

```
***

# Problem 4  
## Part 1

```{r}
#huber loss function part 1: "original" function

huber <- function(x)
{
  
  if (abs(x) < 1)
  {
    
    return( x^2 )
    
  } else {
    
    return( 2*abs(x) - 1 )
  
  }

}
#tests
print(huber(1))
print(huber(4))

```
This code should be self explanatory.

## Part 2

```{r}
huber <- function(x,a)
{
  
  if (abs(x) <= a)
  {
    
    return( x^2 )
    
  } else {
    
    return( 2*a*abs(x) - a^2 )
  
  }

}

print(huber(3,2))
print(huber(3,4))
```
Also self explanatory.


## Part 3

```{r}
huber <- function(x,a=1)
{
  
  if (abs(x) <= a)
  {
    
    return( x^2 )
    
  } else {
    
    return( 2*a*abs(x) - a^2 )
  
  }

}

print(huber(3))
```

## Part 4

```{r}
print(huber(a = 1,x = 3))
print(huber(1,3))
```

These return two different values because the variables are named a and x, so if we define a and x in the function call, we merely change the print order.

## Part 5

```{r}
huber <- function(x,a=1)
{
  n = length(x)
  out = ifelse(abs(x) <= rep(a,n), x*x, 2*a*abs(x) - a^2 )
  return(out)
  
}

print(huber(x = 1:6, a = 3))
```

Vectorized results as expected.

***

# Problem 5

```{r}
#define variables
A = matrix(0*1:6, nrow = 2)
b = rep(1,2)
c = rep(2,3)

Ab = matrix(c(A[1,]*b[1], A[2,]*b[2]), nrow = 2)
Ac = matrix(c(A[,1]*c[1], A[,2]*c[2],A[,3]*c[3]), nrow = 3)

```

```{r}
#a times c
AstarC = A*c
```

What follows is the function for standard user inputs.

```{r}
#this kind of product is often called the hadamard product
RowHadamard <- function(n,m,A,b)
{
  M = matrix(to_vec(for(i in 1:n) for(j in 1:m) A[i,j]*b[i]), nrow = n)
  return(M)
}
```

```{r}
DotHadamard <- function(n,m,A,c)
{
  M = matrix(to_vec(for(i in 1:n) for(j in 1:m) A[i,j]*c[j]), nrow = n)
  return(M)
}
```

R treats matrices as a (row) vector of vectors, where each element of the vector is a column.


